---
title: "Reproducible Research Course Project 1"
author: "Rajkumar Ningkonsin"
date: "April 4, 2019"
output: 
  html_document :
    keep_md: true
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Loading and preprocessing the data

### Load the Activity Monitor data into a dataframe and exclude NA values

```{r}
activity <- read.csv("~/Coursera/data/activity.csv")
activity_sub <- na.exclude(activity)
```
### Mean of number of steps per day
``` {r}
num_steps <- aggregate(activity_sub$steps, by = list(date = activity_sub$date), FUN = sum )
```
              
                     
### Histogram of the total number of steps taken each day
```{r plot1}
hist(num_steps$x, main="Total number of steps per day for each date",xlab="Number of Steps", col="darkmagenta")
```


### Summary of total number of steps taken each day 
``` {r echo=TRUE}
steps_summary <- summary(num_steps$x)

steps_summary
```

### Mean of steps taken each day
```{r}
steps_summary["Mean"]
```

### Median of steps taken each day
```{r}
steps_summary["Median"]
```


# What is the average daily activity pattern?

### Time series plot of 5-minute interval and average number of steps taken, averaged across all days 

```{r plot2}
ave_steps <- aggregate(activity_sub$steps, by = list(interval = activity_sub$interval), FUN = mean )

plot(ave_steps$interval, ave_steps$x, type = "l", main = "Average number of steps taken, averaged across all days", ylab = "Number of steps", xlab = "5 min interval", col="darkmagenta")
```

### 5-minute interval with maximum number of steps

``` {r}
subset(ave_steps$interval, ave_steps$x == max(ave_steps$x))
```


# Inputing missing values

### Total number of NAs in the dataset
``` {r}
table(is.na(activity$steps))["TRUE"]
```

### Filling in all of the missing values in the dataset with correspnding mean
``` {r}
activity_na <- activity

for (i in seq_along(activity_na$steps)) {
  if(is.na(activity_na$steps)[i] == TRUE){
    activity_na$steps[i] <- subset(ave_steps$x, ave_steps$interval == activity_na$interval[i])
  }
  else activity_na$steps[i] <- activity_na$steps[i]
}
```
### New dataset equal to the original dataset but with missing data filled in (sample 5 records)
```{r}
head(activity_na,5)
```

### Histogram of total number of steps taken each day with filled-in data
``` {r plot3}
num_steps_na <- aggregate(activity_na$steps, by = list(date = activity_na$date), FUN = sum )

hist(num_steps_na$x, main="Total number of steps per day for each date with NA replaced as mean",xlab="Number of Steps", col="darkmagenta")
```

### Summary of total number of steps taken each day for filled-in data
``` {r echo=TRUE}
steps_summary_na <- summary(num_steps_na$x)

steps_summary_na
```

### Mean of steps taken each day for filled-in data
```{r}
steps_summary_na["Mean"]
```

### Median of steps taken each day for filled-in data
```{r}
steps_summary_na["Median"]
```

# Are there differences in activity patterns between weekdays and weekends?

### Dataset with two levels "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r}
activity_na$date <- as.Date(activity_na$date)

activity_na$day <- weekdays(activity_na$date)

for (i in seq_along(activity_na$day)) {
  if(activity_na$day[i] %in% c("Saturday","Sunday")){
    activity_na$weekend[i] <- "weekend"
  }
  else activity_na$weekend[i] <- "weekday"
}
```

### Dataset with two levels "weekday" and "weekend" (sample 5 records)
```{r}
head(activity_na,5)
```

### Dataset of average steps average across all days by interval and weekend

```{r include=FALSE}
library(dplyr)
```

```{r}
ave_steps_days <- activity_na %>% group_by(interval, weekend) %>% summarise_at(vars(steps), mean)
```


### Dataset of average steps average across all days by interval and weekend (sample 6 records)
```{r}
head(ave_steps_days,6)
```

### Panel plot containing a time series plots of 5-minute interval and average number of steps taken, averaged across all weekday days or weekend days.

```{r plot4}
library(lattice)

xyplot(steps~interval | factor(weekend), data = ave_steps_days, main="Steps vs Interval for weekend vs weekday", xlab="5 min Interval",  ylab="Number of Steps",layout=c(1,2),type="l")
```






